---
title: "Peer Assessment 1 for Reproducible Research"
author: "Spencer Green"
date: "March 5, 2017"
output: html_document
---

Removes any variables in the environment from previous work.

```{r setup}
rm(list=ls()) 
```

Loading and Preprocessing the data.
===================================

We can also do some simple exploratory analysis of what type of data is contained in activity.csv.

```{r echo=TRUE}
dat <- read.csv('activity.csv', header = TRUE)
names(dat)
head(dat)
dim(dat)
str(dat)
```

We can subset the data by day to make later calculations easier. We'll remove missing values at this time.

```{r echo=TRUE}
dat_steps <- aggregate(steps~date,data = dat, FUN = sum, na.rm=TRUE)
```

Mean and Median steps per day.
================================

```{r echo=TRUE}
mean_steps <- mean(dat_steps$steps)
med_steps <- median(dat_steps$steps)
```

The Mean is `r mean_steps` and the Median is `r med_steps`

Now the histogram itself.
  
```{r echo=TRUE}
hist(dat_steps$steps, main="Total Steps per Day", xlab = "Steps Taken in a day", breaks = 20)
```


Time series plot of average steps a day.
========================================

```{r echo=TRUE}
interval_steps <- aggregate(steps~interval, data = dat, FUN = mean, na.rm = TRUE)
plot(steps~interval, data = interval_steps, type = "l")
```

Which 5-minute interval that, on average, contains the maximum number of steps.
===============================================================================

```{r}
interval_steps[which.max(interval_steps$steps),]$interval
```

The most steps is the **`r interval_steps[which.max(interval_steps$steps),]$interval`th** interval.

Describe and show strategy for imputing NAs.
============================================

```{r echo=TRUE}
sum(is.na(dat$steps))
```

The total rows missing are **`r sum(is.na(dat$steps))`**  

Create a new dataset that is equal to the original dataset but with the missing data filled in.  

Strategy: Use the mean of the 5-minute interval.  

First I made a function missed_steps to get the mean steps for particular 5-minute interval. 

```{r echo=TRUE}
missed_steps<-function(interval){
    stepsInterval[stepsInterval$interval==interval,]$steps
}
```

Create the new dataset

```{r echo=TRUE}
complete_steps <- merge(dat, dat_steps, by = "date", suffixes = c("", ".mean"))
nas <- is.na(complete_steps$steps)
complete_steps$steps[nas] <- complete_steps$mean[nas]
complete_steps <- complete_steps[, c(1:3)]
head(complete_steps)
```

Histogram of the total steps taken each day:

```{r echo=TRUE}
dat_steps <- aggregate(steps ~ date, complete_steps, sum)
barplot(dat_steps$steps, names.arg = dat_steps$date, ylim = c(0,25000),xlab = "date",ylab = "sum(steps)",)
```

Are There Differences between Weekdays and Weekends in Activity?
================================================================

First, we need to differentiate weekend days and weekdays  

```{r echo=TRUE}
complete_steps$day=ifelse(as.POSIXlt(as.Date(complete_steps$date))$wday%%6==0,"weekend","weekday")
complete_steps$day=factor(complete_steps$day,levels = c("weekday","weekend"))
head(complete_steps)
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken

```{r echo=TRUE}
steps_interval <- aggregate(steps~interval+day,complete_steps,mean)
library(lattice)
xyplot(steps~interval | factor(day), data=steps_interval, aspects=1/2, type="l")
```

As we can see, there's generally more activity over weekends, but weekdays have a larger morning spike.
